### 2009-04-03 talk/icooolps2009-dotnet/intro.tex
Try to adapt section 1.2 to avoid repeating from the Introduction.

The alternative is to write a compiler; writing a compiler that targets a high
level virtual machine like CLI or JVM is easier than targeting a real CPU, but
it still requires a lot of work, as IronPython, Jython, JRuby demonstrate.

Moreover, writing a static compiler is often not enough to get high
performance; IronPython and JRuby are going in the direction of JIT compiling
specialized versions of the code depending on the actual values/types seen at
runtime; this approach seems to work, but writing it manually requires an
enormous effort.

@@

The first step is to write an interpreter for the chosen language.  Since it
must be fed to the translation toolchain, the interpreter has to be written in
RPython.  Then, to guide the process, we need to add few manual
annotations to the interpreter, in order to teach the JIT generator which
informations are important to know at compile-time.  Annotations are inserted
as \emph{hints}, as described in section \ref{sec:hints}.

@@

Generating efficient compilers for dynamic languages is hard.  Since these
languages are dynamically typed, usually the compiler does not have enough
information to produce efficient code, but instead it has to insert a lot of
runtime checks to select the appropriate implementation for each operation.

By emitting code at runtime, JIT compilers can exploit some extra knowledge
compared to traditional static compilers.  However, we need to take special
care to choose a strategy for JIT compilation that lets the compiler to take
the best of this advantage.

Most JIT compilers for dynamic languages around (such as
IronPython\footnote{http://www.codeplex.com/IronPython},
Jython\footnote{http://www.jython.org/} or
JRuby\footnote{http://jruby.codehaus.org/}) compile code at the method
granularity.  If on the one hand they can exploit some of the knowledge
gathered at runtime (e.g. the types of method parameters), on the other hand
they can do little to optimize most of the operations inside, because few
assumptions can be made about the global state of the program.

JIT compilers generated by PyPy solve this problem by delaying the compilation
until they know all the informations needed to generate efficient code.  If at
some point the JIT compiler does not know about something it needs, it
generates a callback into itself and stops execution.

>>>>

The alternative is to write a compiler; writing a compiler that targets a high
level virtual machine like CLI or JVM is easier than targeting a real CPU, but
it still requires a lot of work, as
IronPython\footnote{http://www.codeplex.com/IronPython},
Jython\footnote{http://www.jython.org/}
and JRuby\footnote{http://jruby.codehaus.org/} demonstrate.

Moreover, writing a static compiler is often not enough to get high
performance.  IronPython and JRuby are going in the direction of
compiling just-in-time (JIT) specialized versions of the code depending
on the actual values/types seen at runtime; this approach seems to work,
but writing it manually requires an enormous effort.

@@

The main difference between the JIT compilers generated by PyPy and the
ones found in other projects like IronPython is that the latter compile
code at the method granularity -- if on the one hand they can exploit
some of the knowledge gathered at runtime (e.g.\ the types of method
parameters), on the other hand they can do little to optimize most of
the operations inside, because few assumptions can be made about the
global state of the program.  The PyPy JITs, on the other hand, work at
a sub-method granularity, as described next.

When using PyPy, the first step is to write an interpreter for the chosen language.  Since it
must be fed to the translation toolchain, the interpreter has to be written in
RPython.  Then, to guide the process, we need to add few manual
annotations to the interpreter, in order to teach the JIT generator which
information is important to know at compile-time.  Annotations are inserted
as \emph{hints}, as described in section \ref{sec:hints}.
From these hints, PyPy will statically generate an interpreter and a JIT
compiler in a single executable (here a .NET executable).

@@

\commentout{
By emitting code at runtime, JIT compilers can exploit extra
knowledge compared to traditional static compilers -- more than just the
type of the function arguments.  However, special care is needed to
choose a strategy for JIT compilation that lets the compiler take the
best of this advantage.
}

The interesting property of the generated JIT compiler is to delay the
compilation until it knows all the information needed to generate
efficient code.  In other words, at runtime, when the interpreter notice
that it is useful to compile a given piece of code, it sends it to the
JIT compiler; however, if at some point the JIT compiler does not know
about something it needs, it generates a callback into itself and stops
execution.
